Efficiently stable presentations from error-correcting codes

### 2023-07-09 efficient_stability.tex
some additional comments in Section 5

\begin{claim}\label{claim:z2-stab-2b}
For $u\in \F_q^m$, $j\in\{1,\ldots,m\}$ and $i\in \{0,\ldots,d\}$ let $U_{i} = U_{u+ie_j}$ and, for all $\beta\in\F_q$, $P^{(i)}_\beta = P^{u+ie_j}_\beta$ (leaving the dependence on $u$ implicit). Then 
\begin{equation}\label{eq:z2-stab-2b-0a}
\Es{u\in \F_q^m} \Es{\substack{j\in\{1,\ldots,m\}\\i\neq i' \in \{0,\ldots,d\}}}\Es{\alpha,\alpha'\in \F_q} \big\| \big[ U_i^\alpha, U_{i'}^{\alpha'}\big]-\Id\big\|_\tau^2\,=\, O\big(\sqrt{\eps}\big)\;, 
\end{equation}
and
\begin{equation}\label{eq:z2-stab-2b-0b}
 \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}}\Es{\alpha,\alpha'\in \F_q} \big\| \big[ U_u^\alpha, U_{v}^{\alpha'}\big]-\Id\big\|_\tau^2\,=\, O\big(\sqrt{\eps}\big)\;,
\end{equation}
where the expectation is over a uniformly random $u$ and $j$, and a uniformly random $v$ conditioned on its last $(j-1)$ coordinates matching those of $u$. 
\end{claim}

\begin{proof}
Due to the test of the relations $\{R^\com_k\}$, it holds that 
\begin{align}
\Es{u\in \F_q^m} \Es{\substack{j\in\{1,\ldots,m\} \\ i\neq i' \in \{0,\ldots,d\}}} \Es{\alpha,\beta\in \F_q} \big\| [\phi(s_{u+i e_j,\alpha}),\phi(s_{u+i' e_j,\beta})] - \Id \big\|_\tau^2&\leq 8\eps\; ,\label{eq:z2-stab-2b-1}\\
\Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}}\Es{\alpha,\alpha'\in \F_q} \big\| [\phi(s_{u,\alpha}),\phi(s_{v,\beta})]-\Id\big\|_\tau^2&\leq 8\eps\;.\label{eq:z2-stab-2b-1b}
\end{align}
Now, for any $u,v\in \F_q^m$
\begin{align*}
\Es{\alpha,\alpha'\in \F_q} &\tau\big( U_u^\alpha U_{v}^{\alpha'} U_u^{-\alpha} U_{v}^{-\alpha'} \big) \\
&= \sum_{\beta,\beta',\gamma,\gamma'\in \F_q}\Es{\alpha\in \F_q} \omega_q^{\alpha(\beta-\beta')} \Es{\alpha'\in \F_q} \omega_q^{\alpha'(\gamma-\gamma')}  \tau(P^{u}_\beta P^{v}_\gamma P^{u}_{\beta'} P^{v}_{\gamma'}\big)\\
&=\sum_{\beta,\beta',\gamma,\gamma'\in \F_q}\Es{\alpha\in\F_2^t} (-1)^{\alpha\cdot\kappa(\beta-\beta')} \Es{\alpha'\in\F_2^t} (-1)^{\alpha'\cdot\kappa(\gamma-\gamma')} \tau(P^{u}_\beta P^{v}_\gamma P^{u}_{\beta'} P^{v}_{\gamma'}\big)\\
&=\Es{\alpha,\alpha'\in\F_2^t}  \tau\Big( \Big(\sum_{\beta\in \F_q} (-1)^{\alpha\cdot \kappa(\beta)} P^{u}_\beta\Big) \Big(\sum_{\gamma\in \F_q} (-1)^{\alpha'\cdot \kappa(\gamma)} P^{v}_\gamma\Big)\Big(\sum_{\beta'\in \F_q} (-1)^{\alpha\cdot \kappa(\beta')} P^{u}_{\beta'}\Big)\Big(\sum_{\gamma'\in \F_q} (-1)^{\alpha'\cdot \kappa(\gamma')} P^{v}_{\gamma'}\Big)\Big)\;.
\end{align*}
Here, the second line follows since for $\beta,\beta'\in \F_q$,
\[ \Es{\alpha\in \F_q} \omega_q^{\alpha(\beta-\beta')}\,=\,\Es{\alpha\in\F_2^t} (-1)^{\alpha\cdot \kappa(\beta-\beta')}\,=\delta_{\beta,\beta'}\;.\]
It follows that, for any distribution on $(u,v)$ such that both marginals are uniform over $\F_q^m$,
\begin{align*}
\Es{u,v} \Es{\alpha,\alpha'\in \F_q} \big\| [U_u^\alpha ,U_{v}^{\alpha'} ]-\Id\big\|_\tau^2
&= \Es{\alpha,\alpha'\in \F_2^t} \Big\| \Big[\sum_{\beta\in \F_q} (-1)^{\alpha\cdot \kappa(\beta)} P^{u}_\beta ,\sum_{\beta'\in \F_2^t} (-1)^{\alpha\cdot \kappa(\beta')} P^{v}_{\beta'} \Big]-\Id\Big\|_\tau^2\\
&= \Es{\alpha,\alpha'\in \F_2^t} \big\| \big[ \phi(s_{u,\alpha}),\phi(s_{v,\alpha'})\big]-\Id\big\|_\tau^2+O\big(\sqrt{\eps}\big)\;,
\end{align*}
where the second line follows from Claim~\ref{claim:z2-stab-2}. Thus~\eqref{eq:z2-stab-2b-0a} follows from~\eqref{eq:z2-stab-2b-1}, and~\eqref{eq:z2-stab-2b-0b} follows from~\eqref{eq:z2-stab-2b-1b}.
\end{proof}

@@

\begin{claim}\label{claim:z2-stab-3}
For every $u\in \F_q^m$ and $j\in\{1,\ldots,m\}$, there is a projective measurement $\{R^{u,j}_\alpha\}_{\alpha\in\F_q^{d+1}}$ on $\mM$ such that 
\[ \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{x_0,\ldots,x_{d}\in \F_q} \Big\| U_0^{x_0}\cdots U_d^{x_d} -  \sum_{\alpha} \omega^{\sum x_i\alpha_i} R^{u,j}_\alpha\Big\|_\tau^2 \,=\, \poly(d,\eps)\;.\]
Similarly, for any $u,v\in \F_q^m$ there is a projective measurement $\{R^{u,v}_{\alpha,\beta}\}_{(\alpha,\beta)\in\F_q^{2}}$ on $\mM$ such that 
\[  \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}} \Es{x,y\in \F_q} \Big\| U_u^{x}  U_v^{y} -  \sum_{\alpha} \omega^{ \alpha x + \beta y} R^{u,v}_{\alpha,\beta}\Big\|_\tau^2 \,=\, \poly(\eps)\;.\]
\end{claim}

>>>>

\begin{claim}\label{claim:z2-stab-2b}
For $u\in \F_q^m$, $j\in\{1,\ldots,m\}$ and $i\in \{0,\ldots,d\}$ let $U_{i} = U_{u+ie_j}$ and, for all $\beta\in\F_q$, $P^{(i)}_\beta = P^{u+ie_j}_\beta$ (leaving the dependence on $u$ \hnote{and $j$?} implicit). Then 
\begin{equation}\label{eq:z2-stab-2b-0a}
\Es{u\in \F_q^m} \Es{\substack{j\in\{1,\ldots,m\}\\i\neq i' \in \{0,\ldots,d\}}}\Es{\alpha,\alpha'\in \F_q} \big\| \big[ U_i^\alpha, U_{i'}^{\alpha'}\big]-\Id\big\|_\tau^2\,=\, O\big(\sqrt{\eps}\big)\;, 
\end{equation}
and
\begin{equation}\label{eq:z2-stab-2b-0b}
 \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}}\Es{\alpha,\alpha'\in \F_q} \big\| \big[ U_u^\alpha, U_{v}^{\alpha'}\big]-\Id\big\|_\tau^2\,=\, O\big(\sqrt{\eps}\big)\;,
\end{equation}
where the expectation is over a uniformly random $u$ and $j$, and a uniformly random $v$ conditioned on its last $(j-1)$ coordinates matching those of $u$. 
\hnote{should we clarify what it means to raise a unitary to a power in a finite field? I assume there's some canonical identification of elements of $\F_q$ with $\{0,1,2,\ldots,q-1\}$?} \hnote{Should specify this is the group commutator?}
\end{claim}

\begin{proof}
Due to the test of the relations $\{R^\com_k\}$, it holds that 
\begin{align}
\Es{u\in \F_q^m} \Es{\substack{j\in\{1,\ldots,m\} \\ i\neq i' \in \{0,\ldots,d\}}} \Es{\alpha,\beta\in \F_q} \big\| [\phi(s_{u+i e_j,\alpha}),\phi(s_{u+i' e_j,\beta})] - \Id \big\|_\tau^2&\leq 8\eps\; ,\label{eq:z2-stab-2b-1}\\
\Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}}\Es{\alpha,\alpha'\in \F_q} \big\| [\phi(s_{u,\alpha}),\phi(s_{v,\beta})]-\Id\big\|_\tau^2&\leq 8\eps\;.\label{eq:z2-stab-2b-1b}
\end{align}
Now, for any $u,v\in \F_q^m$
\begin{align*}
\Es{\alpha,\alpha'\in \F_q} &\tau\big( U_u^\alpha U_{v}^{\alpha'} U_u^{-\alpha} U_{v}^{-\alpha'} \big) \\
&= \sum_{\beta,\beta',\gamma,\gamma'\in \F_q}\Es{\alpha\in \F_q} \omega_q^{\alpha(\beta-\beta')} \Es{\alpha'\in \F_q} \omega_q^{\alpha'(\gamma-\gamma')}  \tau(P^{u}_\beta P^{v}_\gamma P^{u}_{\beta'} P^{v}_{\gamma'}\big)\\
&=\sum_{\beta,\beta',\gamma,\gamma'\in \F_q}\Es{\alpha\in\F_2^t} (-1)^{\alpha\cdot\kappa(\beta-\beta')} \Es{\alpha'\in\F_2^t} (-1)^{\alpha'\cdot\kappa(\gamma-\gamma')} \tau(P^{u}_\beta P^{v}_\gamma P^{u}_{\beta'} P^{v}_{\gamma'}\big)\\
&=\Es{\alpha,\alpha'\in\F_2^t}  \tau\Big( \Big(\sum_{\beta\in \F_q} (-1)^{\alpha\cdot \kappa(\beta)} P^{u}_\beta\Big) \Big(\sum_{\gamma\in \F_q} (-1)^{\alpha'\cdot \kappa(\gamma)} P^{v}_\gamma\Big)\Big(\sum_{\beta'\in \F_q} (-1)^{\alpha\cdot \kappa(\beta')} P^{u}_{\beta'}\Big)\Big(\sum_{\gamma'\in \F_q} (-1)^{\alpha'\cdot \kappa(\gamma')} P^{v}_{\gamma'}\Big)\Big)\;.
\end{align*}
Here, the second line follows since for $\beta,\beta'\in \F_q$, \hnote{rearranged so that the delta was in the middle}
\[ \Es{\alpha\in \F_q} \omega_q^{\alpha(\beta-\beta')}\,=\delta_{\beta,\beta'}\,=\,\Es{\alpha\in\F_2^t} (-1)^{\alpha\cdot \kappa(\beta-\beta')}\,\;.\]
It follows that, for any distribution on $(u,v)$ such that both marginals are uniform over $\F_q^m$,
\begin{align*}
\Es{u,v} \Es{\alpha,\alpha'\in \F_q} \big\| [U_u^\alpha ,U_{v}^{\alpha'} ]-\Id\big\|_\tau^2
&= \Es{\alpha,\alpha'\in \F_2^t} \Big\| \Big[\sum_{\beta\in \F_q} (-1)^{\alpha\cdot \kappa(\beta)} P^{u}_\beta ,\sum_{\beta'\in \F_2^t} (-1)^{\alpha\cdot \kappa(\beta')} P^{v}_{\beta'} \Big]-\Id\Big\|_\tau^2\\
&= \Es{\alpha,\alpha'\in \F_2^t} \big\| \big[ \phi(s_{u,\alpha}),\phi(s_{v,\alpha'})\big]-\Id\big\|_\tau^2+O\big(\sqrt{\eps}\big)\;,
\end{align*}
where the second line follows from Claim~\ref{claim:z2-stab-2}. Thus~\eqref{eq:z2-stab-2b-0a} follows from~\eqref{eq:z2-stab-2b-1}, and~\eqref{eq:z2-stab-2b-0b} follows from~\eqref{eq:z2-stab-2b-1b}.
\end{proof}

@@

\begin{claim}\label{claim:z2-stab-3}
For every $u\in \F_q^m$ and $j\in\{1,\ldots,m\}$, there is a projective measurement $\{R^{u,j}_\alpha\}_{\alpha\in\F_q^{d+1}}$ on $\mM$ such that 
\[ \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{x_0,\ldots,x_{d}\in \F_q} \Big\| U_0^{x_0}\cdots U_d^{x_d} -  \sum_{\alpha} \omega^{\sum x_i\alpha_i} R^{u,j}_\alpha\Big\|_\tau^2 \,=\, \poly(d,\eps)\;.\]
\hnote{the use of $x$'s as elements in $\F_q$ is confusing, especially since above we used $x_i$'s as generators...}
Similarly, for any $u,v\in \F_q^m$ there is a projective measurement $\{R^{u,v}_{\alpha,\beta}\}_{(\alpha,\beta)\in\F_q^{2}}$ on $\mM$ such that 
\[  \Es{u\in \F_q^m} \Es{j\in\{1,\ldots,m\}} \Es{\substack{v\in \F_q^m \\v_{m-j+2}=u_{m-j+2},\ldots,v_m=u_m}} \Es{x,y\in \F_q} \Big\| U_u^{x}  U_v^{y} -  \sum_{\alpha} \omega^{ \alpha x + \beta y} R^{u,v}_{\alpha,\beta}\Big\|_\tau^2 \,=\, \poly(\eps)\;.\]
\end{claim}
